Master thesis

### 2017-07-24 docs/master-thesis.tex
Final soutenance

Our work is based on the general hypothesis of a positive relationships
between biodiversity and productivity. We assumed that when a
disturbance event happen, due to a higher productivity, forest with an
increased diversity will recover quicker and thus be more resilient.
This theoritical expectation stand on two processes: complementarity and
selection effects \citep{Loreau2001}. If we take a species pool with
different productivity in monoculture. Their assemblage will allow theim
an overall higher productivity due to a better ressources acquisition
through ressources partitionning and niche differentiation.
Additionally, some species will individually have an higher productivity
in the assemblage than in monoculture due to facilitation from other
species present in the assemblage. Complementarity effect is the
addition of the better ressource acquisition and the facilitation. Now
if we look at the evolution of the species assemblage over time, more
competitive species will progressively dominate through competitive
selection. And if competitive species are more productive, they will
increase assemblage overall productivity. This ressources preemption by
more competitive species is the selection effect. Moreover, we expect
the sampling effect to improve more diverse forest. A bigger initial
sampling of the regional species pool in a rich forest, will allow more
redundancy and a lower risk to lose important functional traits
assemblages for the ecosystem when the disturbance happen.

@@

TROLL model each tree indivdually in a located environment. Thus TROLL
model, alongside with SORTIE \citep{Pacala1996, Uriarte2009} and FORMIND
\citep{Fischer2016, Kohler1998}, can be defined as an individual-based
and spatially explicit forest growth model. TROLL simulates the life
cycle of individual trees from recruitment, with a diameter at breast
height (dbh) above 1 cm, to death with growth and seed production. Trees
are growing in a spatialized light environment explicitly computed witin
voxels of 1 \(m^3\). Each tree is consistently defined by its age,
diameter at brest height (dbh), height (h), crown radius (CR), crown
depth (CD) and leaf area (LA) (see figure \ref{fig:TROLLtree}). Tree
geometry is calculated with allometric equations but leaf area vary
dinamically within each crown following carbon allocations. Voxels
resolution of 1 \(m^3\) allow the establishment of maximum one tree by
1x1 m pixels. Each tree is flagged with a species label inherited from
the parent tree through the seedling recruitment. A species label is
associated to a number of species specific parameters (see table
\ref{tab:traits}) related to functional trait values which can be
sampled on the field.

@@

The underestimation of leaf lifespan for low LMA species with the
allometry from \citet{Reich1991a} resulted in indivduals unealistic
early death from carbon starvation. We gathered data from
TRY\citep{Kattge2011}, DRYAD \citep{chave_towards_2009} and GLOPNET
\citep{wright_worldwide_2004} datasets. We used an out of the bag method
applied on a random forest to select variables with highest importance
to explain leaf lifespan. We thus selected leaf mass per are \(LMA\),
leaf nitrogen content \(N\) and wood specific gravity \(wsg\). We then
used a bayesian approach to test different models with growing level of
complexity. The model with the best tradeoff between complexity (number
of parameters), convergence, likelihood, and prediction quality (root
mean square error of prediction RMSEP) was kept. We selected following
model with a maximum likelihood of 13.6 and a RMSEP of 12 months:

@@

In order to assess the role of biodiversity in ecosystem answer to both
disturbance and sylviculture, we needed to create a space of experiments
encompassing both variation of disturbance, biodiversity and time.
Disturbance was represented by percentage of basal area loss (0\%, 25\%,
50\% and 75\%), or as a selective logging simulation using default
parameters. Biodiversity was integrated with two of its components:
taxonomic and functional diversities. We used species richness \(SR\) to
represents taxonomic diversity (5, 25, and 125 species). Functional
diversity can be related to numerous components, and \citet{Borgy2017}
argued for 5: richness, divergence, regularity, overlap and mean.
Because mature forest were created from a bare soil with TROLL
simulations, we could not control a priori divergence, regularity and
overlap but only assess them after running the simulations, i.e.~before
applying the diturbance. Consequently, we focused on functional richness
with convex hull volume \(CHV\) and functional mean with community
weighted mean \(CWM\). For each level of species richness \(SR\), we
selected 20 communities with growing convex hull volume \(CHV\) but with
a community weighted means close to the regional species pool community
weighted means. Effectivelly, we did not wanted drastic change in
community means that could have more effect than functional richness
itself. This design of experiments resulted in 60 communities
(\(5~SR*20~CHV\)) and 240 simulations
(\(60 ~communities*4~levels~of~disturbance\)) over 600 years (maturity
being assumed after 500 years of regeneration \citep{Li}). Figure
\ref{fig:DOE} presents the design of experiment for communities
biodiversity after the mature forest were simulated, and thus before
disturbance. We obtained a broad range of both functionl dispersion
\(FDis\) and aboveground biomass \(AGB\) for simulated forest ecosystems
before disturbance.

@@

Tropical forest ecosystems provides numerous ecosystem services linked
to several ecosystem functions. We decided to describe simiulated
tropical forests in two major functions: forest structure and forest
functionning. Forest structure was represented by aboveground biomass
(\(AGB\) in \(ton~C.ha^{-1}\)), basal area (\(BA\) in \(m^2.ha^{-1}\)),
total number of stem (\(N\)), number of stem above \(10~cm\) diameter
(\(N10\)), and number of stem above \(30~cm\) diameter (\(N30\)). Forest
functionning was represented by growth pimary productivity (\(GPP\) in
\(MgC.ha^{-1}\)), net pimary productivity (\(NPP\) in \(MgC.ha^{-1}\)),
tree autotrophic respiration in day (\(Rday\) in \(MgC.ha^{-1}\)) and
tree autotrophic respiration in night (\(Rnight\) in \(MgC.ha^{-1}\)).

@@

Biodiversity net effect \(NE\) is based on the difference between
ecosystem variable \(X\) observed value \(X_O\) within the community
mixture of species and its expected value \(X_E\) if species performance
were equal to their performance in monocultures. This effect can be
partitioned between complementarity effect \(CE\), representing niche
partitionning, positive interactions, and resource supply, and selectvie
effect \(SE\) due to dominant species pool driving the ecosystem. Both
metrics depend on the variation of relative ecosystem variable
\(\Delta RX\):

@@

\(X_{sp}\) is the ecosystem variable value for one species either in
mixture \(X_{sp}(mixture)\) or in monoculture \(X_{sp}(monoculture)\).
\(P_{sp}\) is the proportion of the species in the mixture represented
by species relative abundance. Consequently, \(CE\) averages diversity
effects of all species presents in the mixture (both negatives and
positives). Whereas \(SE\) become positive when dominant species
outperform theimselves in mixture than in monoculture, and negative when
less dominant species outperform theimselves in mixture than in
monoculture \citep{Tobner2016}. But similarly to resilience measurement,
biodiversity net effect \(NE\) is a dynamic equilibrium and vary over
time without disturbance. So in order to correctly assess selection and
complementarity effect in answer to disturbance, we normalized it by
undisturbed control ecosystem net effect \(NE_C\) to measure treatments
net effect resilience \(R(NE_T)\):

@@

We measured all ecosystem outputs biodiversity net effect in disturbance
simulations by comparing theim to their species corresponding
monoculture simulations. The net effect was then partition between
selection and complementarity effect. We normalized complementarity and
selection effect of disturbed simulations by biodiversity net effect of
undisturbed control simulation (see Table \ref{tab:tNE}) to measure the
resilience to disturbance of their aboveground biomass (see Figure
\ref{fig:gBE}). We found that complementarity effect was recovering
biodiversity net effect and was stronger than selection effect in the
first decades. But after few decades the complementarity effect
diminsihed toward a low value. On the contrary selection effect was
reduced or even removed by the disturbance. It increased during the
whole simulation and was greater than complementarity effect only after
decades. The time lag for which complementarity effect was greater than
selection effect was increasing with disturbance intensity. Finally 600
years after the disturbance event biodiversity net effect was still not
recovered for a disturbance intensity greater than 25\% of basal area.
We obtained those results for aboveground biomass. We found similar
results but with an aplified signal for basal area (\(BA\)) and stem
abundance (\(N\) but with an inverted signal because of the forest self
thinning) (see
\protect\hyperlink{appendix-5-disturbance-simulations}{Appendix 5:
Disturbance simulations}). Finally, forest growth primary productivity
(\(GPP\)) recovered in few years (proportionnaly to disturbance), and
its net effect was maintained by complementarity effect (see
\protect\hyperlink{appendix-5-disturbance-simulations}{Appendix 5:
Disturbance simulations}).

@@

We found that complementary effect was insuring forest resilience in the
beginning of forest successions. We interpreted this results as the
consequence of facilitation processes. As the only ressource simulated
by TROLL is the light, the main facilitation will be light shading of
post-pioneer species by pioneer species in disturbance gaps. Our results
confirm the study of \citet{Morin2011} who also highlighted the
importance of facilitation through light shading in forest resilience.
But the complementarity effect is reducing through time, to let the
selection effect insure forest resilience due to an inforced dominance
of more productive species. The diminishing of complementarity effect is
due to competitive selection though time in gaps succession. But the
study scale matters as shwon by \citet{Chisholm2013}. Here we look at
processes to a 16 ha scale. We do not have topography reducing
micro-environment effects, but \citet{Chisholm2013} results suggest that
complementarity will be stronger at smaller scale and could explain its
low value after forest recovery. Finally, our findings confirm results
of \citet{Tobner2016} realized on experimental forests of 4 years in low
diverse forest of Canada. They also found that selection effect was
greater than complementarity in most of the cases. In the case of
selective logging the complementarity effect will thus be the major
effect between two cutting cycles due to the cycle length (several
decades in guyana shield). High forest diversity with important species
complementarity through increased functional diversity is thus an
advantage for forest recovery between two cutting cycles in order to
maintain both productive ecosystem and sustainable management.

>>>>

Our work is based on the general hypothesis of a positive relationships
between biodiversity and productivity. We assumed that when a
disturbance event happen, due to a higher productivity, forest with an
increased diversity will recover quicker and thus be more resilient.
This theoritical expectation stand on two processes: complementarity and
selection effects \citep{Loreau2001}. If we take a species pool with
different productivity in monoculture. Their assemblage will allow theim
an overall higher productivity due to a better ressources acquisition
through ressources partitionning and niche differentiation.
Additionnally, some species will individually have an higher
productivity in the assemblage than in monoculture due to facilitation
from other species present in the assemblage. Complementarity effect is
the addition of the better ressource acquisition and the facilitation.
Now if we look at the evolution of the species assemblage over time,
more competitive species will progressively dominate through competitive
selection. And if competitive species are more productive, they will
increase assemblage overall productivity. This ressources preemption by
more competitive species is the selection effect. Moreover, we expect
the sampling effect to improve more diverse forest. A bigger initial
sampling of the regional species pool in a rich forest, will allow more
redundancy and a lower risk to lose important functional traits
assemblages for the ecosystem when the disturbance happen.

@@

TROLL model each tree indivdually in a located environment. Thus TROLL
model, alongside with SORTIE \citep{Pacala1996, Uriarte2009} and FORMIND
\citep{Fischer2016, Kohler1998}, can be defined as an individual-based
and spatially explicit forest growth model. TROLL simulates the life
cycle of individual trees from recruitment, with a diameter at breast
height (dbh) above 1 cm, to death with growth and seed production. Trees
are growing in a spatialized light environment explicitly computed witin
voxels of 1 \(m^3\). Each tree is consistently defined by its age,
diameter at brest height (dbh), height (h), crown radius (CR), crown
depth (CD) and leaf area (LA) (see figure \ref{fig:TROLLtree}). Tree
geometry is calculated with allometric equations but leaf area vary
dynamically within each crown following carbon allocations. Voxels
resolution of 1 \(m^3\) allow the establishment of maximum one tree by
1x1 m pixels. Each tree is flagged with a species label inherited from
the parent tree through the seedling recruitment. A species label is
associated to a number of species specific parameters (see table
\ref{tab:traits}) related to functional trait values which can be
sampled on the field.

@@

The underestimation of leaf lifespan for low LMA species with the
allometry from \citet{Reich1991a} resulted in indivduals unealistic
early death from carbon starvation. We gathered data from
TRY\citep{Kattge2011}, DRYAD \citep{chave_towards_2009} and GLOPNET
\citep{wright_worldwide_2004} datasets. We used an out of the bag method
applied on a random forest to select variables with highest importance
to explain leaf lifespan. We thus selected leaf mass per area \(LMA\),
leaf nitrogen content \(N\) and wood specific gravity \(wsg\). We then
used a bayesian approach to test different models with growing level of
complexity. The model with the best tradeoff between complexity (number
of parameters), convergence, likelihood, and prediction quality (root
mean square error of prediction RMSEP) was kept. We selected following
model with a maximum likelihood of 13.6 and a RMSEP of 12 months:

@@

In order to assess the role of biodiversity in ecosystem answer to both
disturbance and sylviculture, we needed to create a space of experiments
encompassing both variation of disturbance, biodiversity and time.
Disturbance was represented by percentage of basal area loss (0\%, 25\%,
50\% and 75\%), or as a selective logging simulation using default
parameters. Biodiversity was integrated with two of its components:
taxonomic and functional diversities. We used species richness \(SR\) to
represents taxonomic diversity (5, 25, and 125 species). Functional
diversity can be related to numerous components, and \citet{Borgy2017}
argued for 5: richness, divergence, regularity, overlap and mean.
Because mature forest were created from a bare soil with TROLL
simulations, we could not control a priori divergence, regularity and
overlap but only assess them after running the simulations, i.e.~before
applying the diturbance. Consequently, we focused on functional richness
with convex hull volume \(CHV\) and functional mean with community
weighted mean \(CWM\). For each level of species richness \(SR\), we
selected 20 communities with growing convex hull volume \(CHV\) but with
a community weighted means close to the regional species pool community
weighted means. Effectivelly, we did not wanted drastic change in
community means that could have more effect than functional richness
itself. This design of experiments resulted in 60 communities
(\(5~SR*20~CHV\)) and 240 simulations
(\(60 ~communities*4~levels~of~disturbance\)) over 600 years (maturity
being assumed after 500 years of regeneration \citep{Li}). Functional
diversities of mature forests were assessed with
\citet{villeger_new_2008} indices (FRIC, FEve, FDiv, and FDis). Figure
\ref{fig:DOE} presents the design of experiment for communities
biodiversity after the mature forest were simulated, and thus before
disturbance. We obtained a broad range of both functionl dispersion
\(FDis\) and aboveground biomass \(AGB\) for simulated forest ecosystems
before disturbance.

@@

Tropical forest ecosystems provides numerous ecosystem services linked
to several ecosystem functions. We decided to describe simulated
tropical forests in two major functions: forest structure and forest
functionning. Forest structure was represented by aboveground biomass
(\(AGB\) in \(ton~C.ha^{-1}\)), basal area (\(BA\) in \(m^2.ha^{-1}\)),
total number of stem (\(N\)), number of stem above \(10~cm\) diameter
(\(N10\)), and number of stem above \(30~cm\) diameter (\(N30\)). Forest
functionning was represented by growth primary productivity (\(GPP\) in
\(MgC.ha^{-1}\)), net primary productivity (\(NPP\) in \(MgC.ha^{-1}\)),
tree autotrophic respiration in day (\(Rday\) in \(MgC.ha^{-1}\)) and
tree autotrophic respiration in night (\(Rnight\) in \(MgC.ha^{-1}\)).

@@

Biodiversity net effect \(NE\) is based on the difference between
ecosystem variable \(X\) observed value \(X_O\) within the community
mixture of species and its expected value \(X_E\) if species performance
were equal to their performance in monocultures. This effect can be
partitioned between complementarity effect \(CE\), representing niche
partitionning, positive interactions, and resource supply, and selectvie
effect \(SE\) due to dominant species pool driving the ecosystem. \(N\)
represents the total number of species, and \(M\) the vector of
monocultures performance. Both metrics depend on the variation of
relative ecosystem variable \(\Delta RX\):

@@

\(X_{sp}\) is the ecosystem variable value for one species either in
mixture \(X_{sp}(mixture)\) or in monoculture \(X_{sp}(monoculture)\).
\(P_{sp}\) is the proportion of the species in the mixture represented
by species relative abundance. Consequently, \(CE\) averages diversity
effects of all species presents in the mixture (both negatives and
positives). Whereas \(SE\) become positive when dominant species
outperform themselves in mixture than in monoculture, and negative when
less dominant species outperform themselves in mixture than in
monoculture \citep{Tobner2016}. But similarly to resilience measurement,
biodiversity net effect \(NE\) is a dynamic equilibrium and vary over
time without disturbance. So in order to correctly assess selection and
complementarity effect in answer to disturbance, we normalized it by
undisturbed control ecosystem net effect \(NE_C\) to measure treatments
net effect resilience \(R(NE_T)\):

@@

We measured all ecosystem outputs biodiversity net effect in disturbance
simulations by comparing them to their species corresponding monoculture
simulations. The net effect was then partition between selection and
complementarity effect. We normalized complementarity and selection
effect of disturbed simulations by biodiversity net effect of
undisturbed control simulation (see Table \ref{tab:tNE}) to measure the
resilience to disturbance of their aboveground biomass (see Figure
\ref{fig:gBE}). We found that complementarity effect was recovering
biodiversity net effect and was stronger than selection effect in the
first decades. But after few decades the complementarity effect
diminsihed toward a low value. On the contrary selection effect was
reduced or even removed by the disturbance. It increased during the
whole simulation and was greater than complementarity effect only after
decades. The time lag for which complementarity effect was greater than
selection effect was increasing with disturbance intensity. Finally 600
years after the disturbance event biodiversity net effect was still not
recovered for a disturbance intensity greater than 25\% of basal area.
We obtained those results for aboveground biomass. We found similar
results but with an amplified signal for basal area (\(BA\)) and stem
abundance (\(N\) but with an inverted signal because of the forest self
thinning) (see
\protect\hyperlink{appendix-5-disturbance-simulations}{Appendix 5:
Disturbance simulations}). Finally, forest growth primary productivity
(\(GPP\)) recovered in few years (proportionnaly to disturbance), and
its net effect was maintained by complementarity effect (see
\protect\hyperlink{appendix-5-disturbance-simulations}{Appendix 5:
Disturbance simulations}).

@@

We found that complementary effect was insuring forest resilience in the
beginning of forest successions. We interpreted this results as the
consequence of facilitation processes. As the only ressource simulated
by TROLL is the light, the main facilitation will be light shading of
post-pioneer species by pioneer species in disturbance gaps. Our results
confirm the study of \citet{Morin2011} who also highlighted the
importance of facilitation through light shading in forest resilience.
But the complementarity effect is reducing through time, to let the
selection effect insure forest resilience due to an inforced dominance
of more productive species. The diminishing of complementarity effect is
due to competitive selection though time in gaps succession. But the
study scale matters as shown by \citet{Chisholm2013}. Here we look at
processes to a 16 ha scale. We do not have topography reducing
micro-environment effects, but \citet{Chisholm2013} results suggest that
complementarity will be stronger at smaller scale and could explain its
low value after forest recovery. Finally, our findings confirm results
of \citet{Tobner2016} realized on experimental forests of 4 years in low
diverse forest of Canada. They also found that selection effect was
greater than complementarity in most of the cases. In the case of
selective logging the complementarity effect will thus be the major
effect between two cutting cycles due to the cycle length (several
decades in guyana shield). High forest diversity with important species
complementarity through increased functional diversity is thus an
advantage for forest recovery between two cutting cycles in order to
maintain both productive ecosystem and sustainable management.
